### 2023-04-18 scripts/ms_scaling_changes.qmd
initial commit

---
title: "Decomposing biodiversity change to processes of extinction, colonization, and persistence across scales"
author: "François Leroy et al."
format: 
  html:
    toc: true
    toc_float: true
editor: visual
---

This document contains the code to reprodcuce the models fitting and figures for our manuscript. Models were created for the 2 datasets we used, *i.e.* the Czech bird Atlases (hereafter CzAtlas) and the Czech breeding bird survey (hereafter CzBBS). Those datasets are provided already processed (*i.e.* species richness, colonization, extinction... per unit of area). If you are interested in accessing those datasets, feel free to contact the following email address: leroy\@fzp.czu.cz.

# Model fitting

## Species richness

### CzAtlas

```{r, message=FALSE, warning=FALSE}
rm(list = ls())
library(caret)
library(caretEnsemble)
library(dplyr)
library(ranger)
library(ggtext)
library(cowplot)
library(forcats)
library(tidyr)
library(stringr)
```

```{r}
## Load data
load("../data/data_model_fitting/atlas_cart_data.Rdata")

## We perform repeated (n = 3) cross-validation (k = 10) using the caret package
# First we split the dataset into training and validation datasets
# set.seed(333)
traindata <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

```

```{r, eval=FALSE}
## We use three tree-based algorithms in order to assess which one perform the best
## Create the list of models
model_list_atlas <-
  caretList(
    sr ~ ., 
    data = atlas_cart_data,
    trControl = traindata,
    metric = "RMSE",
    tuneList = list(
      rf =  caretModelSpec(method = "ranger", 
                           importance = "permutation",
                           tuneGrid = data.frame(mtry = ceiling((ncol(atlas_cart_data)-1)/3),
                                                                    splitrule = "variance",
                                                                    min.node.size = 5)),
      gbm = caretModelSpec(method = "gbm",
                           distribution = "poisson",
                           # keep.data = TRUE,
                           tuneGrid = data.frame(n.trees = 500,
                                                 interaction.depth = 1,
                                                 shrinkage = 0.1,
                                                 n.minobsinnode = 5)),
      xgboost = caretModelSpec(method = "xgbTree",
                               tuneGrid = expand.grid(nrounds = 500,
                                                     max_depth = 5,
                                                     eta = 0.1,
                                                     gamma = 0,
                                                     colsample_bytree = 1,
                                                     min_child_weight = 1,
                                                     subsample = 1))
    )
  )

# saveRDS(model_list_atlas, "../models/sr_atlas.rds")
```

### CzBBS (or JPSP)

```{r}
# Load data
load("../data/data_model_fitting/JPSP_cart_data.Rdata")

# First, convert the date into a numeric format (here time stamp since 1970-01-01)
JPSP_cart_data$Date <- as.numeric(JPSP_cart_data$Date)

## Reduce the time_span
JPSP_cart_data <- JPSP_cart_data %>% dplyr::filter(time_span <= 5 & time_span >= 3)
```

```{r, eval=FALSE}
## Create the list of models
model_list_JPSP <-
  caretList(
    sr ~ ., 
    data = JPSP_cart_data,
    trControl = traindata,
    metric = "RMSE",
    tuneList = list(
      rf =  caretModelSpec(method = "ranger",
                           importance = "permutation",
                           tuneGrid = data.frame(mtry = ceiling((ncol(JPSP_cart_data)-1)/3),
                                                                    splitrule = "variance",
                                                                    min.node.size = 5)),
      gbm = caretModelSpec(method = "gbm",
                           distribution = "poisson",
                           tuneGrid = data.frame(n.trees = 500,
                                                 interaction.depth = 1,
                                                 shrinkage = 0.1,
                                                 n.minobsinnode = 5)),
      xgboost = caretModelSpec(method = "xgbTree",
                               tuneGrid = expand.grid(nrounds = 500,
                                                     max_depth = 5,
                                                     eta = 0.1,
                                                     gamma = 0,
                                                     colsample_bytree = 1,
                                                     min_child_weight = 1,
                                                     subsample = 1))
    )
  )

# saveRDS(model_list_JPSP, file = "../models/sr_bbs.rds")
```

## Number of Colonization, Extinction and Persistence (CEP)

### CzAtlas

```{r}
## Load data for learning
CEP_atlas_cart_data <- readRDS("../data/data_model_fitting/CEP_atlas_cart_data.rds")
CEP_atlas_cart_data <- CEP_atlas_cart_data[!grepl("^proba", CEP_atlas_cart_data$metric),]
```

```{r, eval=FALSE}
## For each metric, perform repeated cross-validation with the 3 algorithms
## Learn for each metric
for(THE_metric in unique(CEP_atlas_cart_data$metric)){

  assign(
    paste0("CEP_atlas_models_", THE_metric),
    caretList(
      value ~ ., 
      data = CEP_atlas_cart_data %>% filter(metric == THE_metric) %>% dplyr::select(-quadrat, -metric, -trials),
      trControl = traindata,
      metric = "RMSE",
      tuneList = list(
        rf =  caretModelSpec(method = "ranger",
                             importance = "permutation",
                             tuneGrid = data.frame(mtry = ceiling((ncol(CEP_atlas_cart_data)-2)/3),
                                                                      splitrule = "variance",
                                                                      min.node.size = 5)),
        gbm = caretModelSpec(method = "gbm",
                             tuneGrid = data.frame(n.trees = 500,
                                                   interaction.depth = 1,
                                                   shrinkage = 0.1,
                                                   n.minobsinnode = 5)),
        xgboost = caretModelSpec(method = "xgbTree",
                                 tuneGrid = expand.grid(nrounds = 500,
                                                       max_depth = 5,
                                                       eta = 0.1,
                                                       gamma = 0,
                                                       colsample_bytree = 1,
                                                       min_child_weight = 1,
                                                       subsample = 1))
      )
    )
  )
}
### Merge all of them together
CEP_models <- list()
CEP_models[["col"]] <-  CEP_atlas_models_col
CEP_models[["ext"]] <-  CEP_atlas_models_ext
CEP_models[["pers"]] <-  CEP_atlas_models_pers
saveRDS(CEP_models, file = "../models/nCEP_atlas.rds")
```

### CzBBS

```{r}
CEP_JPSP_cart_data <- readRDS("../data/data_model_fitting/CEP_JPSP_cart_data.rds")
CEP_JPSP_cart_data <- CEP_JPSP_cart_data[!grepl("^proba", CEP_JPSP_cart_data$metric),]
```

```{r, eval=FALSE}
for(THE_metric in unique(CEP_JPSP_cart_data$metric)){

  assign(
    paste0("CEP_JPSP_models_", THE_metric),
    caretList(
      value ~ ., 
      data = CEP_JPSP_cart_data %>% 
        filter(metric == THE_metric) %>%
        dplyr::select(-metric) %>% 
        filter(!is.na(value)),
      trControl = traindata,
      metric = "RMSE",
      tuneList = list(
        rf =  caretModelSpec(method = "ranger",
                             importance = "permutation",
                             tuneGrid = data.frame(mtry = ceiling((ncol(CEP_JPSP_cart_data)-2)/3),
                                                                      splitrule = "variance",
                                                                      min.node.size = 5)),
        gbm = caretModelSpec(method = "gbm",
                             tuneGrid = data.frame(n.trees = 500,
                                                   interaction.depth = 1,
                                                   shrinkage = 0.1,
                                                   n.minobsinnode = 5)),
        xgboost = caretModelSpec(method = "xgbTree",
                                 tuneGrid = expand.grid(nrounds = 500,
                                                       max_depth = 5,
                                                       eta = 0.1,
                                                       gamma = 0,
                                                       colsample_bytree = 1,
                                                       min_child_weight = 1,
                                                       subsample = 1))
      )
    )
  )
}
### Merge all of them together
CEP_models_JPSP <- list()
CEP_models_JPSP[["col"]] <-  `CEP_JPSP_models_colonization`
CEP_models_JPSP[["ext"]] <-  `CEP_JPSP_models_extinction`
CEP_models_JPSP[["pers"]] <-  `CEP_JPSP_models_persistence`
saveRDS(CEP_models_JPSP, file = "../models/nCEP_JPSP.rds")
```

## Probability of Colonization, Extinction and Persitence

### CzAtlas

```{r}
## Load data for learning
CEP_atlas_cart_data <- readRDS("../data/data_model_fitting/CEP_atlas_cart_data.rds")
CEP_atlas_cart_data <- CEP_atlas_cart_data[grepl("^proba", CEP_atlas_cart_data$metric),]
```

```{r, eval=FALSE}
## For each metric, perform repeated cross-validation with the 3 algorithms
## Learn for each metric
for(THE_metric in unique(CEP_atlas_cart_data$metric)){

  assign(
    paste0("CEP_atlas_models_", THE_metric),
    caretList(
      value ~ ., 
      data = CEP_atlas_cart_data %>% filter(metric == THE_metric) %>% dplyr::select(-quadrat, -metric, -trials),
      trControl = traindata,
      metric = "RMSE",
      tuneList = list(
        rf =  caretModelSpec(method = "ranger",
                             importance = "permutation",
                             tuneGrid = data.frame(mtry = ceiling((ncol(CEP_atlas_cart_data)-2)/3),
                                                                      splitrule = "variance",
                                                                      min.node.size = 5)),
        gbm = caretModelSpec(method = "gbm",
                             tuneGrid = data.frame(n.trees = 500,
                                                   interaction.depth = 1,
                                                   shrinkage = 0.1,
                                                   n.minobsinnode = 5)),
        xgboost = caretModelSpec(method = "xgbTree",
                                 tuneGrid = expand.grid(nrounds = 500,
                                                       max_depth = 5,
                                                       eta = 0.1,
                                                       gamma = 0,
                                                       colsample_bytree = 1,
                                                       min_child_weight = 1,
                                                       subsample = 1))
      )
    )
  )
}
### Merge all of them together
CEP_models <- list()
CEP_models[["col"]] <-  CEP_atlas_models_proba_colonized
CEP_models[["ext"]] <-  CEP_atlas_models_proba_ext
CEP_models[["pers"]] <-  CEP_atlas_models_proba_pers
saveRDS(CEP_models, file = "../models/probaCEP_atlas.rds")
```

### CzBBS

```{r}
CEP_JPSP_cart_data <- readRDS("../data/data_model_fitting/CEP_JPSP_cart_data.rds")
CEP_JPSP_cart_data <- CEP_JPSP_cart_data[grepl("^proba", CEP_JPSP_cart_data$metric),]
```

```{r, eval=FALSE}
for(THE_metric in unique(CEP_JPSP_cart_data$metric)){

  assign(
    paste0("CEP_JPSP_models_", THE_metric),
    caretList(
      value ~ ., 
      data = CEP_JPSP_cart_data %>% 
        filter(metric == THE_metric) %>%
        dplyr::select(-metric) %>% 
        filter(!is.na(value)),
      trControl = traindata,
      metric = "RMSE",
      tuneList = list(
        rf =  caretModelSpec(method = "ranger",
                             importance = "permutation",
                             tuneGrid = data.frame(mtry = ceiling((ncol(CEP_JPSP_cart_data)-2)/3),
                                                                      splitrule = "variance",
                                                                      min.node.size = 5)),
        gbm = caretModelSpec(method = "gbm",
                             tuneGrid = data.frame(n.trees = 500,
                                                   interaction.depth = 1,
                                                   shrinkage = 0.1,
                                                   n.minobsinnode = 5)),
        xgboost = caretModelSpec(method = "xgbTree",
                                 tuneGrid = expand.grid(nrounds = 500,
                                                       max_depth = 5,
                                                       eta = 0.1,
                                                       gamma = 0,
                                                       colsample_bytree = 1,
                                                       min_child_weight = 1,
                                                       subsample = 1))
      )
    )
  )
}
### Merge all of them together
CEP_models_JPSP <- list()
CEP_models_JPSP[["col"]] <-  `CEP_JPSP_models_proba colonized`
CEP_models_JPSP[["ext"]] <-  `CEP_JPSP_models_proba extinction`
CEP_models_JPSP[["pers"]] <-  `CEP_JPSP_models_proba persistence`
saveRDS(CEP_models_JPSP, file = "../models/probaCEP_JPSP.rds")
```

# Predictions

## Species richness

```{r}
# Load the models
models_list_atlas <- readRDS(file = "../models/sr_atlas.rds")
models_list_JPSP <- readRDS(file = "../models/sr_bbs.rds")
```

### CzAtlas

```{r}
## Load the data used for predictions
load("../data/data_to_predict/atlas_data_topredict.Rdata")
```

```{r}
## In order to set a fixed values of sampling effort for the different spatial scales, we fit a quadratic regression between area and sampling effort
polynome <- lm(effort ~ poly(spatial_scale, 2), data = atlas_data_topredict)
```

```{r}
## Original grid cell ######
pred130sqkm <-
  atlas_data_topredict %>%
  filter(spatial_scale == 130) %>% 
  dplyr::select(-c(KVADRAT, sr, time_span, effort, spatial_scale)) %>% 
  mutate(time_span = 4,
         effort = round(predict(polynome, newdata = data.frame(spatial_scale = 130))))
## Make the prediction
pred130sqkm$pred_sr <- round(predict(models_list_atlas$xgboost$finalModel, newdata = as.matrix(pred130sqkm)))
pred130sqkm$KVADRAT <- atlas_data_topredict$KVADRAT[atlas_data_topredict$spatial_scale == 130]
## And compute the trend
### M2 --> M4 
slope_130_M2M4 <- 
pred130sqkm %>% 
  group_by(KVADRAT) %>% 
  do({
    mod <-  lm(pred_sr ~ Date, data = .)
    data.frame(slope = coef(mod)[2])
  })
### M2 --> M3 
slope_130_M2M3 <- 
pred130sqkm %>% 
  filter(Date != 2015) %>% 
  group_by(KVADRAT) %>% 
  do({
    mod <-  lm(pred_sr ~ Date, data = .)
    data.frame(slope = coef(mod)[2])
  })
### M3 --> M4
slope_130_M3M4 <- 
pred130sqkm %>% 
  filter(Date != 1987) %>% 
  group_by(KVADRAT) %>% 
  do({
    mod <-  lm(pred_sr ~ Date, data = .)
    data.frame(slope = coef(mod)[2])
  })

## 2 by 2 grids ######
## Prepare the dataset
pred530sqkm <-
  atlas_data_topredict %>%
  filter(spatial_scale == 530) %>% 
  dplyr::select(-c(KVADRAT, sr, time_span, effort, spatial_scale)) %>% 
  mutate(time_span = 4,
         effort = round(predict(polynome, newdata = data.frame(spatial_scale = 530))))
## Make the prediction
pred530sqkm$pred_sr <- round(predict(models_list_atlas$xgboost$finalModel, newdata = as.matrix(pred530sqkm)))
pred530sqkm$KVADRAT <- atlas_data_topredict$KVADRAT[atlas_data_topredict$spatial_scale == 530]
## And compute the trend
### M2 --> M4
slope_530_M2M4 <-
pred530sqkm %>% 
  group_by(KVADRAT) %>% 
  do({
    mod <-  lm(pred_sr ~ Date, data = .)
    data.frame(slope = coef(mod)[2])
  }) %>% 
  ## take off the cells of different area which contains less than 4 subcells
  filter_at(vars(KVADRAT), any_vars(grepl("^[^,\n]*((,[^,\n]*){3}$)", ., perl = T)))
### M2 --> M3
slope_530_M2M3 <-
pred530sqkm %>% 
  filter(Date != 2015) %>% 
  group_by(KVADRAT) %>% 
  do({
    mod <-  lm(pred_sr ~ Date, data = .)
    data.frame(slope = coef(mod)[2])
  }) %>% 
  ## take off the cells of different area which contains less than 4 subcells
  filter_at(vars(KVADRAT), any_vars(grepl("^[^,\n]*((,[^,\n]*){3}$)", ., perl = T)))
### M3 --> M4
slope_530_M3M4 <-
pred530sqkm %>% 
  filter(Date != 1987) %>% 
  group_by(KVADRAT) %>% 
  do({
    mod <-  lm(pred_sr ~ Date, data = .)
    data.frame(slope = coef(mod)[2])
  }) %>% 
  ## take off the cells of different area which contains less than 4 subcells
  filter_at(vars(KVADRAT), any_vars(grepl("^[^,\n]*((,[^,\n]*){3}$)", ., perl = T)))

## 4 by 4 cells ######
## Prepare the dataset
pred2150sqkm <-
  atlas_data_topredict %>%
  filter(spatial_scale == 2150) %>% 
  dplyr::select(-c(KVADRAT, sr, time_span, effort, spatial_scale)) %>% 
  mutate(time_span = 4,
         effort = round(predict(polynome, newdata = data.frame(spatial_scale = 2150))))
## Make the prediction
pred2150sqkm$pred_sr <- round(predict(models_list_atlas$xgboost$finalModel, newdata = as.matrix(pred2150sqkm)))
pred2150sqkm$KVADRAT <- atlas_data_topredict$KVADRAT[atlas_data_topredict$spatial_scale == 2150]
## And compute the trend
## M2 --> M4 
slope_2150_M2M4 <-
pred2150sqkm %>% 
  group_by(KVADRAT) %>% 
  do({
    mod <-  lm(pred_sr ~ Date, data = .)
    data.frame(slope = coef(mod)[2])
  }) %>% 
  filter_at(vars(KVADRAT), any_vars(grepl("^[^,\n]*((,[^,\n]*){15}$)", ., perl = T)))
## M2 --> M3
slope_2150_M2M3 <-
pred2150sqkm %>% 
  filter(Date != 2015) %>% 
  group_by(KVADRAT) %>% 
  do({
    mod <-  lm(pred_sr ~ Date, data = .)
    data.frame(slope = coef(mod)[2])
  }) %>% 
  filter_at(vars(KVADRAT), any_vars(grepl("^[^,\n]*((,[^,\n]*){15}$)", ., perl = T)))
## M3 --> M4
slope_2150_M3M4 <-
pred2150sqkm %>% 
  filter(Date != 1987) %>% 
  group_by(KVADRAT) %>% 
  do({
    mod <-  lm(pred_sr ~ Date, data = .)
    data.frame(slope = coef(mod)[2])
  }) %>% 
  filter_at(vars(KVADRAT), any_vars(grepl("^[^,\n]*((,[^,\n]*){15}$)", ., perl = T)))

## Entire Czechia ######
predCR <- 
  atlas_data_topredict %>%
  filter(spatial_scale == 83900) %>% 
  dplyr::select(-c(KVADRAT, sr, time_span, effort, spatial_scale)) %>% 
  mutate(time_span = 4,
         effort = round(predict(polynome, newdata = data.frame(spatial_scale = 83900))))
## Make the prediction
predCR$pred_sr <- round(predict(models_list_atlas$xgboost$finalModel, newdata = as.matrix(predCR)))
predCR$KVADRAT <- atlas_data_topredict$KVADRAT[atlas_data_topredict$spatial_scale == 83900]
## And compute the trend
### M2 --> M4
slope_CR_M2M4 <- 
predCR %>% 
  group_by(KVADRAT) %>% 
  do({
    mod <-  lm(pred_sr ~ Date, data = .)
    data.frame(slope = coef(mod)[2])
  })
### M2 --> M3
slope_CR_M2M3 <- 
predCR %>% 
  filter(Date != 2015) %>% 
  group_by(KVADRAT) %>% 
  do({
    mod <-  lm(pred_sr ~ Date, data = .)
    data.frame(slope = coef(mod)[2])
  })
### M3 --> M4
slope_CR_M3M4 <- 
predCR %>% 
  filter(Date != 1987) %>% 
  group_by(KVADRAT) %>% 
  do({
    mod <-  lm(pred_sr ~ Date, data = .)
    data.frame(slope = coef(mod)[2])
  })
```

### CzBBS

```{r}
## Load data used for predictions
load("../data/data_to_predict/JPSP_topredict.Rdata")
```

```{r, warning=F}
## Create the predictions
JPSP_predictions <- predict(models_list_JPSP$rf$finalModel, data = JPSP_topredict %>% dplyr::select(-id))
## Add them to the data
JPSP_topredict$sr <- round(JPSP_predictions$predictions)
## Now compute the trends
### M2 --> M4
slope_JPSP_M2M4 <- 
JPSP_topredict %>% 
  group_by(id, AREA) %>% 
  do({
    mod <-  lm(sr ~ Date, data = .)
    data.frame(slope = coef(mod)[2],
               pval = (summary(mod)$coefficients[2,"Pr(>|t|)"]))
  }) %>% 
  left_join(JPSP_topredict %>% dplyr::select(id, AREA), by = c("id", "AREA")) %>% 
  unique()
#### M2 --> M3
slope_JPSP_M2M3 <-
JPSP_topredict %>% 
  filter(Date != 2015) %>% 
  group_by(id, AREA) %>% 
  do({
    mod <-  lm(sr ~ Date, data = .)
    data.frame(slope = coef(mod)[2],
               pval = (summary(mod)$coefficients[2,"Pr(>|t|)"]))
  }) %>% 
  left_join(JPSP_topredict %>% dplyr::select(id, AREA), by = c("id", "AREA")) %>% 
  unique()
#### M3 --> M4
slope_JPSP_M3M4 <-
JPSP_topredict %>% 
  filter(Date != 1987) %>% 
  group_by(id, AREA) %>% 
  do({
    mod <-  lm(sr ~ Date, data = .)
    data.frame(slope = coef(mod)[2],
               pval = (summary(mod)$coefficients[2,"Pr(>|t|)"]))
  }) %>% 
  left_join(JPSP_topredict %>% dplyr::select(id, AREA), by = c("id", "AREA")) %>% 
  unique()
```

## Number of Colonization, Extinction and Persistence

### CzAtlas

```{r}
nCEP_atlas_models <- readRDS("../models/nCEP_atlas.rds")
```

```{r}
## Create the dataframe to predict
nCEP_topredict <-
atlas_data_topredict %>% 
  # filter(!(((KVADRAT == 6844 & Date == 2015 & effort != 5))|(KVADRAT == 7248 & Date == 2015 & effort != 5))) %>% 
  filter(AREA != 0) %>% 
  dplyr::select(-c(Date, time_span, sr, effort)) %>% 
  distinct() %>% 
  mutate(effort_diff = 0,
         timespan_diff = 0) %>% 
  ## add the periods
  crossing(period = c("M2M3", "M3M4", "M2M4"))

## And make predictions for each metric
predicted_nCEP_atlas <- data.frame()
## predictions for each metric
for(THE_metric in names(nCEP_atlas_models)){
  model <- nCEP_atlas_models[[THE_metric]][["xgboost"]]
  pred <- predict(model, 
                  newdata = nCEP_topredict)
  predicted_nCEP_atlas <- 
  rbind(predicted_nCEP_atlas,
        nCEP_topredict %>% cbind(pred = round(pred)) %>% mutate(metric = THE_metric))
}
## Also, compute jaccard index
## Compute turnover
predicted_nCEP_atlas <- 
predicted_nCEP_atlas %>% 
  pivot_wider(id_cols = c(Lat, Long, AREA, polypoint_ratio, KVADRAT, elevation, spatial_scale, effort_diff, timespan_diff),
              names_from = c(metric,period),
              values_from = pred,
              names_sep = "") %>% 
  group_by(Lat, Long, AREA, polypoint_ratio, KVADRAT, elevation, spatial_scale, effort_diff, timespan_diff) %>% 
  mutate(jaccardM2M3 = persM2M3/(persM2M3 + colM2M3 + extM2M3),
         jaccardM3M4 = persM3M4/(persM3M4 + colM3M4 + extM3M4),
         jaccardM2M4 = persM2M4/(persM2M4 + colM2M4 + extM2M4)) %>% 
  ungroup()
```

### CzBBS

```{r}
nCEP_BBS_models <- readRDS("../models/nCEP_JPSP.rds")
```

```{r}
predicted_nCEP_BBS <- data.frame()
for(THE_metric in names(nCEP_BBS_models)){
  model <- nCEP_BBS_models[[THE_metric]][["rf"]]
  pred <- 
    predict(model, 
            newdata = JPSP_topredict %>% 
              dplyr::select(-time_span, -id, -sr, -Date, polypoint_ratio, Lat, Long, elevation, AREA) %>% 
              mutate(timespan_diff = 0) %>% 
              crossing(period = c("M2M3", "M3M4", "M2M4")))
  predicted_nCEP_BBS <- 
  rbind(
    predicted_nCEP_BBS,
    JPSP_topredict %>% 
    dplyr::select(time_span, Lat, Long, elevation, AREA) %>% 
    mutate(metric = THE_metric) %>% 
    crossing(period = c("M2M3", "M3M4", "M2M4")) %>% 
    cbind(pred = round(pred))
  )
}
## Compute turnover
predicted_nCEP_BBS <- 
predicted_nCEP_BBS %>% 
  pivot_wider(id_cols = c(Lat, Long, AREA, elevation, time_span),
              names_from = c(metric,period),
              values_from = pred,
              names_sep = "") %>% 
  group_by(Lat, Long, AREA, elevation, time_span) %>% 
  mutate(jaccardM2M3 = persM2M3/(persM2M3 + colM2M3 + extM2M3),
         jaccardM3M4 = persM3M4/(persM3M4 + colM3M4 + extM3M4),
         jaccardM2M4 = persM2M4/(persM2M4 + colM2M4 + extM2M4)) %>% 
  ungroup()
```

# Figures

## Species richness

```{r, warning=FALSE, message=FALSE}
# M2M4 ####
pdf("../figures/sr_spatialscaling_M2M4.pdf",
    height = 5, width = 5.5)
# Data management
slope_JPSP_M2M4 %>% 
  dplyr::rename(area = AREA) %>% 
  dplyr::select(slope, area) %>% 
  mutate(area = ifelse(area < 0.2, 0.13, 2.51)) %>% 
  rbind(
    slope_130_M2M4 %>% 
      dplyr::select(slope) %>% 
      mutate(area = 130) %>% 
      rbind(
        slope_530_M2M4 %>% 
          dplyr::select(slope) %>% 
          mutate(area = 530)
      ) %>% 
      rbind(
        slope_2150_M2M4 %>% 
          dplyr::select(slope) %>% 
          mutate(area = 2150)
      ) %>% 
      rbind(
        slope_CR_M2M4 %>% 
          dplyr::select(slope) %>% 
          mutate(area = 83900)
      )
  ) %>% 
  ## Plot
  ggplot() +
  geom_boxplot(aes(area, slope, group = area))+
  scale_x_log10()+
  xlab("Spatial grain (km²)")+
  ylab("Species Richness change\nper year")+
  ylim(-2,2)+
    theme(panel.background = element_blank(),
        panel.grid.major = element_line(colour = "lightgrey"),
        axis.text.x = element_markdown())+
  stat_summary(aes(area, slope), fun.y = median, geom = 'line', linetype = 2, size = 1,  color = "blue")+
  stat_summary(aes(area, slope), fun = mean,
               geom = "point", col = "red", size = 1)+
  stat_summary(aes(area, slope), fun.data = mean_se,
               geom = "errorbar", size = .5, width = .1, color = "red")+
  geom_hline(yintercept = 0)+
  annotation_logticks(sides = "b")+
  coord_cartesian(clip = "off")+
  theme(text = element_text(size = 16))
dev.off()

# M2M3 ####
pdf("../figures/sr_spatialscaling_M2M3.pdf",
    height = 5, width = 5.5)
slope_JPSP_M2M3 %>% 
  dplyr::rename(area = AREA) %>% 
  dplyr::select(slope, area) %>% 
  mutate(area = ifelse(area < 0.2, 0.13, 2.51)) %>% 
  rbind(
    slope_130_M2M3 %>% 
      dplyr::select(slope) %>% 
      mutate(area = 130) %>% 
      rbind(
        slope_530_M2M3 %>% 
          dplyr::select(slope) %>% 
          mutate(area = 530)
      ) %>% 
      rbind(
        slope_2150_M2M3 %>% 
          dplyr::select(slope) %>% 
          mutate(area = 2150)
      ) %>% 
      rbind(
        slope_CR_M2M3 %>% 
          dplyr::select(slope) %>% 
          mutate(area = 83900)
      )
  ) %>% 
  ggplot() +
  geom_boxplot(aes(area, slope, group = area))+
  scale_x_log10()+
  xlab("Spatial grain (km²)")+
  ylab("Species Richness change\nper year")+
  ylim(-2,2)+
    theme(panel.background = element_blank(),
        panel.grid.major = element_line(colour = "lightgrey"),
        axis.text.x = element_markdown())+
  stat_summary(aes(area, slope), fun.y = median, geom = 'line', linetype = 2, size = 1,  color = "blue")+
  stat_summary(aes(area, slope), fun = mean,
               geom = "point", col = "red", size = 1)+
  stat_summary(aes(area, slope), fun.data = mean_se,
               geom = "errorbar", size = .5, width = .1, color = "red")+
  geom_hline(yintercept = 0)+
  annotation_logticks(sides = "b")+
  coord_cartesian(clip = "off")+
  theme(text = element_text(size = 16))
dev.off()
# M3M4 ####
pdf("../figures/sr_spatialscaling_M3M4.pdf",
    height = 5, width = 5.5)
slope_JPSP_M3M4 %>% 
  dplyr::rename(area = AREA) %>% 
  dplyr::select(slope, area) %>% 
  mutate(area = ifelse(area < 0.2, 0.13, 2.51)) %>% 
  rbind(
    slope_130_M3M4 %>% 
      dplyr::select(slope) %>% 
      mutate(area = 130) %>% 
      rbind(
        slope_530_M3M4 %>% 
          dplyr::select(slope) %>% 
          mutate(area = 530)
      ) %>% 
      rbind(
        slope_2150_M3M4 %>% 
          dplyr::select(slope) %>% 
          mutate(area = 2150)
      ) %>% 
      rbind(
        slope_CR_M3M4 %>% 
          dplyr::select(slope) %>% 
          mutate(area = 83900)
      )
  ) %>% 
  ggplot() +
  geom_boxplot(aes(area, slope, group = area))+
  scale_x_log10()+
  xlab("Spatial grain (km²)")+
  ylab("Species Richness change\nper year")+
  ylim(-2,2)+
    theme(panel.background = element_blank(),
        panel.grid.major = element_line(colour = "lightgrey"),
        axis.text.x = element_markdown())+
  stat_summary(aes(area, slope), fun.y = median, geom = 'line', linetype = 2, size = 1,  color = "blue")+
  stat_summary(aes(area, slope), fun = mean,
               geom = "point", col = "red", size = 1)+
  stat_summary(aes(area, slope), fun.data = mean_se,
               geom = "errorbar", size = .5, width = .1, color = "red")+
  geom_hline(yintercept = 0)+
  annotation_logticks(sides = "b")+
  coord_cartesian(clip = "off")+
  theme(text = element_text(size = 16))
dev.off()
## Figure 4
pdf("../figures/fig4_sr.pdf",
    height = 5, width = 5.5)
slope_JPSP_M2M3 %>% 
  dplyr::rename(area = AREA) %>% 
  dplyr::select(slope, area) %>% 
  mutate(area = ifelse(area < 0.2, 0.13, 2.51)) %>% 
  rbind(
    slope_130_M2M3 %>% 
      dplyr::select(slope) %>% 
      mutate(area = 130) %>% 
      rbind(
        slope_530_M2M3 %>% 
          dplyr::select(slope) %>% 
          mutate(area = 530)
      ) %>% 
      rbind(
        slope_2150_M2M3 %>% 
          dplyr::select(slope) %>% 
          mutate(area = 2150)
      ) %>% 
      rbind(
        slope_CR_M2M3 %>% 
          dplyr::select(slope) %>% 
          mutate(area = 83900)
      ) 
  ) %>% mutate(period = "M2M3") %>% 
  rbind(
    slope_JPSP_M3M4 %>% 
      dplyr::rename(area = AREA) %>% 
      dplyr::select(slope, area) %>% 
      mutate(area = ifelse(area < 0.2, 0.13, 2.51)) %>% 
      rbind(
        slope_130_M3M4 %>% 
          dplyr::select(slope) %>% 
          mutate(area = 130) %>% 
          rbind(
            slope_530_M3M4 %>% 
              dplyr::select(slope) %>% 
              mutate(area = 530)
          ) %>% 
          rbind(
            slope_2150_M3M4 %>% 
              dplyr::select(slope) %>% 
              mutate(area = 2150)
          ) %>% 
          rbind(
            slope_CR_M3M4 %>% 
              dplyr::select(slope) %>% 
              mutate(area = 83900)
          ) 
       ) %>% mutate(period = "M3M4")
    ) %>% 
  mutate(period = case_when(
    period == "M2M3" ~ "1985-2003",
    period == "M3M4" ~ "2003-2017"
  )) %>% 
  mutate(period = fct_relevel(period, c("1985-2003", "2003-2017"))) %>% 
  ggplot()+
  # geom_violin(aes(area, slope, group = interaction(area, period)))+
  geom_boxplot(aes(area, slope, group = interaction(area, period), color = period))+
  stat_summary(aes(area, slope, color = period), fun.y = median, geom = 'line', size = 1, linetype = 2)+
  # stat_summary(aes(area, slope, color = period, shape = period), fun.y = median, geom = 'point', size = 2.5)+
  # ylim(-2, 2)+
  scale_x_log10()+
  annotation_logticks(sides = "b")+
  xlab("Spatial grain (km²)")+
  ylab("Species Richness change\nper year")+
  labs(color = "Period", shape = "Period", linetype = "Period")+
  ylim(-2,2)+
  theme_bw()+
  theme(text = element_text(size = 16))+
  theme(panel.background = element_blank(),
        legend.position = "none",
        panel.grid.major = element_line(colour = "lightgrey"),
        axis.text.x = element_markdown())
dev.off()

```

## Number of Colonization, Extinction, Persistence and Turnover

```{r}
pdf("../figures/nCEP_scaling.pdf",
    height = 5, width = 5.5)
for(periods in c("M2M3", "M3M4", "M2M4")){
  print(
    predicted_nCEP_atlas %>% dplyr::select(spatial_scale, contains("M", ignore.case = F)) %>% dplyr::rename(AREA = spatial_scale) %>% 
    rbind(
      predicted_nCEP_BBS %>% dplyr::select(contains("M", ignore.case = F), AREA) %>% 
        mutate(AREA = ifelse(AREA < 0.2, 0.13, 2.51))
    ) %>% 
      pivot_longer(cols = -c("AREA"),
                 names_to = "period",
                 values_to = "value") %>% 
    separate(period, into = c("metric", "period"), sep = c(-4)) %>% 
    filter(metric == "col" | metric == "ext" | metric == "pers") %>% 
  filter(period == periods) %>% 
  ggplot()+
  geom_boxplot(aes(AREA, value,
                   group = interaction(AREA,metric),
                   fill = metric),
                   outlier.shape = NA)+
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")+
  scale_x_log10()+
  scale_y_log10(limits = c(4,210))+
  annotation_logticks(sides = "bl")+
  stat_summary(aes(AREA, value, color = metric), fun.y = median, geom = 'line', linetype = 2, size = 1)+
  ylab("Count")+
  xlab("Spatial grain\n(km²)")+
  labs(fill = "Probability")+
  theme_bw()+
  theme(text = element_text(size = 16),
        legend.position = "none")
  )
}
dev.off()
```

```{r}
pdf("../figures/col_ext_2periods.pdf",
    height = 5, width = 5.5)

for(THE_metric in c("col", "ext")){
print(predicted_nCEP_atlas %>% dplyr::select(spatial_scale, contains("M", ignore.case = F)) %>%
        dplyr::rename(AREA = spatial_scale) %>% 
        rbind(predicted_nCEP_BBS %>% dplyr::select(contains("M", ignore.case = F), AREA) %>% 
                mutate(AREA = ifelse(AREA < 0.2, 0.13, 2.51))) %>% 
        dplyr::select(AREA, contains(THE_metric)) %>% 
        pivot_longer(cols = contains(THE_metric),
                     names_to = "period",
                     values_to = "value") %>% 
        mutate(period = str_remove(period, THE_metric)) %>% 
          mutate(period = case_when(
            period == "M2M4" ~ "1985-2017",
            period == "M2M3" ~ "1985-2003",
            period == "M3M4" ~ "2003-2017")) %>% 
        mutate(period = fct_relevel(period, c("1985-2017", "1985-2003", "2003-2017"))) %>%
        filter(period != "1985-2017") %>%
        ggplot()+
        geom_boxplot(aes(AREA, value, group = interaction(AREA, period), color = period))+
        stat_summary(aes(AREA, value, color = period), fun.y = median, geom = 'line', size = 1, linetype = 2)+
        scale_x_log10()+
        annotation_logticks(sides = "b")+
        xlab("Spatial grain (km²)")+
        ylab(THE_metric)+
        ylim(0,50)+
        labs(color = "Period", shape = "Period", linetype = "Period")+
        theme_bw()+
        theme(text = element_text(size = 16),
              legend.position = "none"))
}
dev.off()
```

```{r}
pdf("../figures/jaccard.pdf",
    height = 5, width = 5.5)

for(period in c("jaccardM2M3","jaccardM3M4")){
print(predicted_nCEP_atlas %>% 
        dplyr::select(spatial_scale, contains("M", ignore.case = F)) %>%
        dplyr::rename(AREA = spatial_scale) %>% 
        rbind(predicted_nCEP_BBS %>% 
        dplyr::select(contains("M", ignore.case = F), AREA) %>% 
        mutate(AREA = ifelse(AREA < 0.2, 0.13, 2.51))) %>% 
        dplyr::select(AREA, period) %>% 
        dplyr::rename(value = period) %>% 
        ggplot(aes(AREA, value, group = AREA))+
        annotation_logticks(sides = "b")+
        geom_boxplot()+
        scale_x_log10()+
        xlab("Spatial grain\n(km²)")+
        ylim(.35, 1)+
        ylab("Jaccard")+
        theme_bw()+
        theme(strip.background = element_blank(),
              strip.text.x = element_blank(),
              text = element_text(size = 16))+
        stat_summary(aes(group = 1), fun.y = median,
                     geom = "line", linetype = "dashed", size = 1, color = "blue"))
}
dev.off()
```